---
title: "Simplex Algorithms"
author: "Alex Kloska"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stargazer)
library(kableExtra)
library(hash)
library(MASS)
options(digits=10,scipen=999)
```

# Create Matrix Function
```{r}
createLP <- function(C,...,varnames=NULL,integers=NULL,basic=NULL){
  x <- list(...)
  M <- as.matrix(C)
  for(i in 1:length(x)){
    M <- cbind(M,x[[i]])
  }
  M <- t(M)
  I <- diag(nrow=nrow(M),ncol=nrow(M))
  I <- apply(I,2,lag)
  I[1,] <- 0
  I <- I[,-ncol(I)]
  RHS <- as.matrix(M[,ncol(M)])
  M <- M[,-ncol(M)]
  M <- cbind(M,I,RHS)
  M[1,] <- -1*M[1,] # making objective row negative
  tab <- cbind(c(1,rep(0,nrow(M)-1)), M)
  tab -> M
  if(is.null(varnames)){
    varnames <- c('z',paste0("x",1:(ncol(M)-length(x)-2)),paste0("s",1:(length(x))),'RHS')
  }
  if(is.null(integers)){
    integers <- rep(F,ncol(M))
  }
  variables <- list(varnames = varnames,
                    integers = integers,
                    basic = basic)
  M <- unname(M)
  lp <- list(matrix = M, variables = variables, identity = I, objVar = (length(C)-1))
  return(lp)
}
```

# Creating Dual
```{r}
dualLP <- function(A){
  c <- A[,ncol(A)]
}
```

# Primal Simplex using Blands Anti-Cycling Algorithm for Strictly Linear Programs
```{r, warning=FALSE, message=FALSE}
simplex <- function(A){
  library(kableExtra)
  # setting column and row names to numeric values
  colnames(A) = c(1:ncol(A))
  rownames(A) = c(1:nrow(A))
  count <- 0
  repeat{
    count <- count + 1
    # Locating column of first negative coefficient in objective function
    for(i in 1:ncol(A)){
      x <- A[2:nrow(A),i]
      if(A[1,i] < 0 & any(x > 0)){ # checking obj coefficient is negative and at least 1 row below it is positive
        which(colnames(A)==i) -> c
        break
      }
    }
    # Dividing entering variable by RHS and selecting row with lowest value
    rvec <- c(NA) # NA so objective function is ignored
    for(i in 2:nrow(A)){
      if(A[i,c] > 0){
        rvec <- append(rvec,(1/((A[i,c] / A[i,ncol(A)])))) # add row to row vector
      } else if(A[i,c] <= 0){
        rvec <- append(rvec,NA) # non-negativity constraint
      }
    }
    if(isTRUE(all.equal(rvec,NA))){ # Entering variable but no leaving variable
      count = 'Unbounded. Ignore Solution'
      break
    }
    which.min(rvec) -> r # select minimum of row vector
    A[r,c] -> pivot # pivot point at r,c
    A[r,] <- (1/pivot)*A[r,] # reduce pivot row
    # row operations to pivot at [r,c]
    for(i in 1:nrow(A)){
      if(rownames(A)[i]==rownames(A)[r]){ # skip pivot row
        next
      }
      A[i,] <- A[i,] - A[i,c]*A[r,]
    }
    # breaks if all coefficients in objective row are 0 or positive
    if(all(A[1,]>=0)){
      break
    }
    if(count==100){ #
      count <- 'Too Many'
      break
    }
  }
  return(A)
}
```

# Primal Simplex Algorithm
```{r message=FALSE, warning=FALSE}
# Step 1: Select a viable entering variable (column)
# Step 2: Select a viable leaving variable (row)
# Step 3: Pivot at [row, column]
# Step 4: Determine if optimal. Return to step 1 if not optimal.

simplex_primal <- function(lp) {
  if(typeof(lp)=="list"){
    A <- lp[[1]]
    if(!is.null(lp[[2]][1])){
      varnames <- lp$variables$varnames
    } else {
      varnames<-c("z",paste0("x",1:(ncol(A)-2)),"RHS")
    }
  } else {
    A <- lp
    varnames <- c('z',paste0("x",1:(ncol(A)-2)),'RHS')
    objVar=1
  }
  if(any(A[2:nrow(A),ncol(A)]<0)){
    stop("Not Initial Basic Primal Feasible. Negative in RHS.")
  }
  if(all(A[1,]>=0)){
    stop("No negative coefficients in cost function.")
  }
  # setting column and row names to numeric values
  colnames(A) = c(1:ncol(A))
  rownames(A) = c(1:nrow(A))
  count <- 0
  repeat{
    count <- count + 1
    # Step 1 -------------------------------------------------------
    # Locating column of first negative coefficient in objective function
    for(i in 1:ncol(A)){
      x <- A[2:nrow(A),i]
      if(A[1,i] < 0 & any(x > 0)){ # checking obj coefficient is negative and at least 1 row below it is positive
        which(colnames(A)==i) -> c
        break
      }
    }
    # Step 2 -------------------------------------------------------
    # Dividing entering variable by RHS and selecting row with lowest value
    rvec <- c(NA) # NA so objective function is ignored
    for(i in 2:nrow(A)){
      if(A[i,c] > 0){
        rvec <- append(rvec,(1/((A[i,c] / A[i,ncol(A)])))) # add row to row vector
      } else if(A[i,c] <= 0){
        rvec <- append(rvec,NA) # non-negativity constraint
      }
    }
    if(isTRUE(all.equal(rvec,NA))){ # Entering variable but no leaving variable
      break
    }
    which.min(rvec) -> r # select minimum of row vector
    # Step 3 -------------------------------------------------------
    print(paste0("Pivot at ",r,",",c))
    A[r,] <- A[r,]/A[r,c] # reduce pivot row
    # row operations to pivot at [r,c]
    for(i in 1:nrow(A)){
      if(rownames(A)[i]==rownames(A)[r]){ # skip pivot row
        next
      }
      A[i,] <- A[i,] - A[i,c]*A[r,]
    }
    # Step 4 -------------------------------------------------------
    # breaks if all coefficients in objective row are 0 or positive
    x <- A[1,]
    if(all(A[1,]>=0)){
      break
    }
    if(count==50){ #
      stop("Too Many Iterations")
    }
  }
  return(A)
}
```

# Kable Implementation
```{r}
kableOut <- function(A){
  decisionVars <- varnames[1:objVar]
  values <- c()
  for(i in 2:(objVar+1)){
    if(sum(A[,i])==1){
      which(colnames(A)==i) -> c
      for(i in 1:nrow(A)){
        if(A[i,c] == 1){
          values <- append(values,A[i,ncol(A)]) 
        } 
      }
    } else {
      values <- append(values,0)
    }
  }
  options(digits=4)
  solution <- data.frame(decisionVars,values)
  # Putting solutions into a string for output display
  solutionOutput <- paste0('Solution:\n')
  for(i in 1:nrow(solution)){
    solutionOutput <- paste0(solutionOutput,solution[i,1],'=',solution[i,2],'\n')
  }
  colnames(A) = varnames
  A %>%
    kable() %>%
    kable_styling(bootstrap_options = c('striped','condensed'), full_width = F, position = 'left') %>%
    footnote(general = paste0('Number of Iterations to Reach Optimum: ',count,'\n',solutionOutput))  -> kabOut
  return(kabOut)
}
```

# Dual Simplex Algorithm
```{r}
# Step 1: Select a viable leaving variable (row)
# Step 2: Select a viable entering variable (column)
# Step 3: Pivot at [row, column]
# Step 4: Determine if optimal. Return to step 1 if not optimal.

simplex_dual <- function(lp){  
  if(typeof(lp)=="list"){
    A <- lp[[1]]
    varnames <- lp$variables$varnames
    identity <- lp[[3]]
  } else {
    A <- lp
  }
  if(!any(A[,ncol(A)]<0)){
    stop("Not Initial Basic Dual Feasible. No negatives in RHS.")
  }
  if(all(A[1,]<0)){
    stop("No positive coefficients in cost function.")
  } 
  # setting column and row names to numeric values
  colnames(A) = c(1:ncol(A))
  rownames(A) = c(1:nrow(A))
  count <- 0
  repeat{
    count <- count + 1
    # Step 1 -------------------------------------------------------
    # Choosing leaving variable
    for(i in 2:nrow(A)){
      x <- A[i,ncol(A)]
      if(A[i,ncol(A)] < 0 & any(x < 0)){
        which(rownames(A)==i) -> r
        break
      } 
    }
    # Step 2 -------------------------------------------------------
    # Dividing entering variable by RHS and selecting row with lowest value
    
    ##### Change to choose first viable column
    cvec <- c() # NA so objective function is ignored
    for(i in 1:(ncol(A)-1)){
      if(A[r,i] < 0){
        cvec <- append(cvec,(A[1,i]/A[r,i])) # add row to row vector
      } else if(A[r,i] >= 0){
        cvec <- append(cvec,NA) # negativity constraint
      }
    }
    if(isTRUE(all.equal(cvec,NA))){ # No leaving variable
      break
    }
    which.max(cvec) -> c # select column closest to zero
    # Step 3 -------------------------------------------------------
    print(paste0("Pivot at ",r,",",c))
    A[r,] <- A[r,]/A[r,c] # reduce pivot row
    # row operations to pivot at [r,c]
    for(i in 1:nrow(A)){
      if(rownames(A)[i]==rownames(A)[r]){ # skip pivot row
        next
      }
      A[i,] <- A[i,] - A[i,c]*A[r,]
    }
    # Step 4 -------------------------------------------------------
    # breaks if all coefficients in objective row are 0 or negative
    if(all(A[2:nrow(A),ncol(A)]>=0)){
      break
    }
    if(count==50){ #
      stop("Too Many Iterations")
    }
  }
  return(A)
}
```

# Sensitivity Analysis
```{r}
# For when the RHS changes
# x is the variable being changed
sensRHS <- function(M, x, delta){
  d <- delta*M[,x] # total change
  M[,ncol(M)] <- M[,ncol(M)] + t(d) # adding to RHS
  M <- cbind(M,d)
  return(M)
}  

# For when the objective function values change
sensCost <- function(M, x, delta){
  r <- match(1,t(M[,x])) # Selecting pivot row
  c <- match(x,colnames(M)) # Selecting pivot column
  M[1,x] <- M[1,x] + delta
  M[r,c] -> pivot # pivot point at r,c
  M[r,] <- (1/pivot)*M[r,] # reduce pivot row
  # row operations to pivot at [r,c]
  for(i in 1:nrow(M)){
    if(rownames(M)[i]==rownames(M)[r]){ # skip pivot row
      next
    }
    M[i,] <- M[i,] - M[i,c]*M[r,]
  }
  return(M)
}
```

# Pivot
```{r,warning=FALSE,message=FALSE}
pivot <- function(A,r,c){
  if(A[r,c]==0){
    stop("Cannot Pivot on 0")
  }
  varnames <- dimnames(A)[[2]]
  colnames(A) = c(1:ncol(A))
  rownames(A) = c(1:nrow(A))
  A[r,] <- A[r,]/A[r,c] # reduce pivot row
  # row operations to pivot at [r,c]
  for(i in 1:nrow(A)){
    if(rownames(A)[i]==rownames(A)[r]){ # skip pivot row
      next
    }
    A[i,] <- A[i,] - A[i,c]*A[r,]
  }
  if(!is.null(varnames)){
    colnames(A)<-varnames
  }
  return(A)
}
```

# Latex Output
```{r}
library(xtable)
latex <- function(A,lengthvec=2){
  print(xtable(A, type = "latex", digits = lengthvec), file = "LatexOutput.tex")
}
```

# Mixed Integer Linear Programming Using Two Phase Simplex
```{r}
# Step 1: Select a viable entering variable (column)
# Step 2: Select a viable leaving variable (row)
# Step 3: Pivot at [row, column]
# Step 4: Check integer. Apply restriction
# Step 5: Determine if optimal. Return to step 1 if not optimal.

simplex_mixedInt <- function(lp) {
  if(typeof(lp)=="list"){
    A <- lp[[1]]
    varnames <- lp[[2]][[1]]
    isInteger <- lp[[2]][[2]]
    basic <- lp[[2]][[3]]
    identity <- lp[[3]]
  } else {
    return(print("Invalid Input"))
  }
  sPrimal <- function(lp) {
    if(typeof(lp)=="list"){
      A <- lp[[1]]
      varnames <- lp[[2]][[1]]
      identity <- lp[[3]]
    } else { 
      A <- lp
      varnames <- c('z',paste0("x",1:(ncol(A)-2)),'RHS')
    }
    library(kableExtra)
    # setting column and row names to numeric values
    colnames(A) = c(1:ncol(A))
    rownames(A) = c(1:nrow(A))
    count <- 0
    repeat{
      count <- count + 1
      # Locating column of first negative coefficient in objective function
      for(i in 1:ncol(A)){
        x <- A[2:nrow(A),i]
        if(A[1,i] < 0 & any(x > 0)){
          which(colnames(A)==i) -> c
          break
        } 
      }
      rvec <- c(NA) # NA so objective function is ignored
      for(i in 2:nrow(A)){
        if(A[i,c] > 0){
          rvec <- append(rvec,(1/((A[i,c] / A[i,ncol(A)])))) # add row to row vector
        } else if(A[i,c] <= 0){ 
          rvec <- append(rvec,NA) # non-negativity constraint
        }
      } 
      if(isTRUE(all.equal(rvec,NA))){ # Entering variable but no leaving variable
        count = 'Unbounded. Ignore Solution'
        break
      }
      which.min(rvec) -> r # select minimum of row vector
      A[r,] <- A[r,]/A[r,c] # reduce pivot row
      # row operations to pivot at [r,c]
      for(i in 1:nrow(A)){
        if(rownames(A)[i]==rownames(A)[r]){ # skip pivot row
          next 
        }
        A[i,] <- A[i,] - A[i,c]*A[r,]
      }
      if(all(t(A[1,]) >= 0)){
        break
      }
      if(count==100){ #
        count <- 'Too Many'
        break
      }
    }
    return(A)
  }
  pivot <- function(A,r,c){
    colnames(A) = c(1:ncol(A))
    rownames(A) = c(1:nrow(A))
    A[r,] <- A[r,]/A[r,c] # reduce pivot row
    # row operations to pivot at [r,c]
    for(i in 1:nrow(A)){
      if(rownames(A)[i]==rownames(A)[r]){ # skip pivot row
        next
      }
      A[i,] <- A[i,] - A[i,c]*A[r,]
    } 
    return(A)
  }
  checkIBFS <- function(A){
    RHS <- A[2:nrow(A),ncol(A)]
    basic <- rep(FALSE,ncol(A))
    for(i in 1:ncol(A)){
      if(identical(unique(A[,i]),c(0,1)) | identical(unique(A[,i]),c(1,0))){
        basic[i] = TRUE
      }
    }
    if(ncol((A[,which(basic==TRUE)])==ncol(diag(nrow(A)))) & all(RHS >= 0)){
      return(TRUE)
    } else {
      return(FALSE)
    }
  }
  library(kableExtra)
  # setting column and row names to numeric values
  colnames(A) = c(1:ncol(A))
  rownames(A) = c(1:nrow(A))
  count <- 0
  # LINEAR PROGRAMMING
  repeat{
    count <- count + 1
    # Step 1 -------------------------------------------------------
    # Locating column of first negative coefficient in objective function
    for(i in 1:ncol(A)){
      x <- A[2:nrow(A),i]
      if(A[1,i] < 0 & any(x > 0)){ # checking obj coefficient is negative and at least 1 row below it is positive
        which(colnames(A)==i) -> c
        break
      }
    }
    # Step 2 -------------------------------------------------------
    # Dividing entering variable by RHS and selecting row with lowest value
    rvec <- c(NA) # NA so objective function is ignored
    for(i in 2:nrow(A)){
      if(A[i,c] > 0){
        rvec <- append(rvec,(1/((A[i,c] / A[i,ncol(A)])))) # add row to row vector
      } else if(A[i,c] <= 0){
        rvec <- append(rvec,NA) # non-negativity constraint
      } 
    } 
    if(isTRUE(all.equal(rvec,NA))){ # Entering variable but no leaving variable
      count = 'Unbounded. Ignore Solution'
      break
    }
    which.min(rvec) -> r # select minimum of row vector
    # Step 3 -------------------------------------------------------
    A[r,] <- A[r,]/A[r,c] # reduce pivot row
    # row operations to pivot at [r,c]
    for(i in 1:nrow(A)){
      if(rownames(A)[i]==rownames(A)[r]){ # skip pivot row
        next
      }
      A[i,] <- A[i,] - A[i,c]*A[r,]
    }
    # Step 4 -------------------------------------------------------
    # Check basis
    for(i in 1:ncol(A)){
      if(identical(unique(A[,i]),c(0,1))){
        basic[i] <- TRUE
      } else {
        basic[i] <- FALSE
      }
    }
    # Step 5 -------------------------------------------------------
    # breaks if all coefficients in objective row are 0 or positive
    if(all(t(A[1,]) >= 0)){
      break
    }
    if(count==100){ #
      count <- 'Too Many'
      break
    }
  }
  # INTEGER PROGRAMMING
  A[1,ncol(A)] -> best
  for(c in 1:(ncol(A)-1)){
    if(isInteger[c]==TRUE & basic[c]==TRUE){
      r <- which(A[,c]==1,arr.ind = TRUE) # Selecting row of basic integer
      if(near(A[r,ncol(A)],as.integer(A[r,ncol(A)]))){
        next
      }
      # Lower Bound
      f <- c(rep(0,ncol(A))) # Creating new vector for new constraint
      f[c] <- 1
      f[ncol(A)] <- floor(A[r,ncol(A)]) # Floor value of basic integer
      Af <- rbind(lp[[1]],f) # Adding vector to original matrix
      Af[,ncol(Af)] -> RHS # Chop RHS to add slack var
      Af <- Af[,-ncol(Af)]
      s1 <- c(rep(0,nrow(Af)-1),1)
      Af <- cbind(Af,s1,RHS) # Floored matrix
      sPrimal(Af) -> Af # Simplex on floored matrix 
      # Grab objective value for comparison 
      if(checkIBFS(Af)==TRUE){
        Af[1,ncol(Af)] -> lb
      } else {
        lb <- 0 # Infeasible
      }
      variables <- list(varnames = varnames,
                        integers = isInteger,
                        basic = basic)
      lpAf <- list(matrix = Af, variables = variables, identity = identity)
      # Upper Bound
      f[ncol(A)] <- ceiling((A[r,ncol(A)])) # Changing constraint to ceiling value of basic integer
      Ac <- rbind(lp[[1]],f) # Row binding f to original matrix
      Ac[,ncol(Ac)] -> RHS # Chop RHS
      e1 <- c(rep(0,nrow(Ac)-1),-1) # Adding excess var
      Ac <- cbind(Ac,e1,RHS)
      objRow <- Ac[1,] # Locking objective row
      a1 <- c(.Machine$integer.max, rep(0,nrow(Ac)-2),1) # Adding artificial var
      #Ac[1,] <- c(1,rep(0,ncol(Ac)-1))
      Ac <- Ac[,-ncol(Ac)]
      Ac <- cbind(Ac,a1,RHS)
      pivot(Ac,nrow(Ac),ncol(Ac)-1) -> Ac
      sPrimal(Ac) -> Ac
      #Ac[,-(ncol(Ac)-1)] -> Ac # Removing artificial variable
      #Ac[1,] <- objRow # Reintroducing objective
      if(checkIBFS(Ac)==TRUE){ # Check to primal
        sPrimal(Ac) -> Ac
      }
      if(checkIBFS(Ac)==TRUE){ # Check to feasible
        Ac[1,ncol(Ac)] -> ub
      } else {
        ub <- 0 # Infeasible
      }
      compVec <- c(best,lb,ub)
      variables <- list(varnames = varnames,
                        integers = isInteger,
                        basic = basic)
      lpAc <- list(matrix = Ac, variables = variables, identity = identity)
      return(list(lpAf=lpAf, lpAc=lpAc))
      #break # needs fixing DO NOT PROCEED
      if(which.max(compVec)==2){ # if lower bound is better, select as next node
        A <- Af
        best <- lb
        if(near(A[r,ncol(A)],as.integer(A[r,ncol(A)]))){
          return(A)
        }
      } else if(which.max(compVec)==3){ # if upper bound is better, select as next node
        A <- Ac
        best <- ub
        if(near(A[r,ncol(A)],as.integer(A[r,ncol(A)]))){
          return(A)
        }
      }
    }
  } 
}
```

# Check Initial Basic Feasible Solution
```{r}
checkIBFS <- function(A){
  RHS <- A[2:nrow(A),ncol(A)]
  basic <- rep(FALSE,ncol(A))
  for(i in 1:ncol(A)){
    if(identical(unique(A[,i]),c(0,1)) | identical(unique(A[,i]),c(1,0))){
      basic[i] = TRUE
    }
  }  
  if(ncol((A[,which(basic==TRUE)])==ncol(diag(nrow(A)))) & all(RHS >= 0)){
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

# Affine Scaling
```{r}
affine <- function(lp,x, alpha=0.9){
  if(typeof(lp)=="list"){
    lp <- lp[[1]]
  } else {
    lp <- lp
  }
  if(missing(x)){
    stop("Initial point x must be specified")
  }
  if(identical(unique(lp[,1]),c(1,0))){
    lp <- lp[,-1] # remove z column if present
  }
  xprev <- x
  repeat{
    C <- -1*lp[1,c(-ncol(lp))]
    M <- lp[-c(1),-c(ncol(lp))]
    D <- diag(x)
    MD <- M%*%D
    CD <- D%*%C
    # (transpose(MD)*inverse(MD*transpose(MD)))*MD
    np <- (t(MD)%*%(ginv(MD%*%t(MD))))%*%MD
    nullProjection <- diag(1,nrow(np),ncol(np)) - np # I - np
    legalDirection <- nullProjection%*%CD
    lambda <- min(legalDirection)
    step <- (-alpha/lambda)*legalDirection
    x <- (x + D%*%step)
    x <- as.vector(x)
    if(any(near(x,0))){
      return(xprev)
    } else if(all(near(x,xprev,tol=.Machine$double.eps))){
      return(x)
    } else {
      xprev <- as.vector(x)
    }
  }
}
```

# Affine Scaling Single Step
```{r}
affineStep <- function(lp,x,alpha=0.9){
  if(typeof(lp)=="list"){
    lp <- lp[[1]]
  } else {
    lp <- lp
  }
  if(missing(x)){
    stop("Initial point x must be specified")
  }
  C <- -1*lp[1,c(-ncol(lp))]
  M <- lp[-c(1),-c(ncol(lp))]
  D <- diag(x)
  MD <- M%*%D
  CD <- D%*%C
  # (transpose(MD)*inverse(MD*transpose(MD)))*MD
  tMD <- t(MD)
  MDtMD <- MD%*%t(MD)
  invMD <- ginv(MDtMD)
  np <- (t(MD)%*%(ginv(MD%*%t(MD))))%*%MD
  nullProjection <- diag(1,nrow(np),ncol(np)) - np # I - np
  legalDirection <- nullProjection%*%CD
  lambda <- min(legalDirection)
  step <- (-alpha/lambda)*legalDirection
  x <- (x + D%*%step)
  x <- as.vector(x)
  return(list(MD=MD,CD=CD,tMD=tMD,MDtMD=MDtMD,invMD=invMD,np=np,nullProjection=nullProjection,legalDirection=legalDirection,step=step,x=x))
}
```

# Add Artificial and Excess Variable
```{r}
addExcessArtificial <- function(M,column,value){
  RHS <- M[,ncol(M)]
  M <- M[,-ncol(M)]
  M <- cbind(M,'e'=c(rep(0,nrow(M))),'a'=c(.Machine$integer.max,rep(0,nrow(M)-1)),RHS)
  M <- rbind(M,c(rep(0,ncol(M))))
  M[nrow(M),ncol(M)-2] <- -1 # excess var
  M[nrow(M),ncol(M)-1] <- 1 # artificial var
  M[nrow(M),column] <- 1
  M[nrow(M),ncol(M)] <- value
  return(M)
}
```

# Check Basis
```{r}
checkBasis <- function(M){
  basis <- c()
  varnames <- c()
  for(i in 1:ncol(M)){
    if(identical(unique(M[,i]),c(0,1)) | identical(unique(M[,i]),c(1,0))){
      r <- which.max(M[,i])
      varnames <- append(varnames,colnames(M)[i])
      basis <- append(basis, M[r,ncol(M)])
    }
  }
  if(length(basis)!=nrow(M)){
    warning(paste("Incomplete Basis. Expected",nrow(M),"basic variables, but only",length(basis),"were found."))
  }
  basis <- setNames(basis,varnames)
  return(basis)
}
```

















